---
title: "Data Experimentation (Pre Report)"
author: "Zehaan Naik"
format: html
editor: visual
---

## CPI Vs SENSEX Data

Loading the data:

```{r}
dat.gold <- read.csv("GoldUP.csv")
CPI.dat <- dat.gold$CPI
SENSEX.dat <- dat.gold$Sensex

CPI.dat <- ts(CPI.dat, start = c(2000,10),  frequency = 12)
SENSEX.dat <- ts(SENSEX.dat, start = c(2000,10),  frequency = 12)
```

Data Summary:

We have monthly data from October 2000 to August 2020.

```{r}
## CPI Data
summary(CPI.dat)

# SENSEX Data
summary(SENSEX.dat)
```

Firstly, let's individually analyse both the covariates as individual time series:

1.  CPI Data:

    ```{r}

    library(tseries)


    # Test for Stationarity
    plot.ts(CPI.dat, ylab = "CPI")
    adf.test(CPI.dat)
    ```

2.  SENSEX Data

    ```{r}

    # Test for Stationarity
    plot.ts(SENSEX.dat, ylab = "SENSEX")
    adf.test(SENSEX.dat)
    ```

As we can see in the case of both our covariates, their individual time series is **not** stationary (by the Augmented Dickey-Fuller Test).

Lastly, before diving into any data transformations or modeling, I wish to try out the Granger test on the two time series:

```{r}
library(lmtest)

grangertest(CPI.dat, SENSEX.dat)
grangertest(SENSEX.dat, CPI.dat)
```

As evident from the Granger test, we can use CPI to predict SENSEX directly however, the same cannot be said about SENSEX serving as a predictor for CPI. This motivates us to consider some transformations on our data to be able to arrive at a better fitting model to us SWAP on.

#### Growth Rate

Firstly, consider a growth rate analysis:

```{r}

# For CPI
CPI.qtrly <- diff(CPI.dat, lag = 4)
plot.ts(CPI.qtrly)
adf.test(CPI.qtrly)

# For SENSEX
SENSEX.qtrly <- diff(SENSEX.dat, lag = 4)
plot.ts(SENSEX.qtrly)
adf.test(SENSEX.qtrly)
```

We observe that differencing at a quaterly time gap does not give us a stationary time series. Hence, analysing the growth rate need not be the best way to go forward.

To further experiment on this we consider a bi-directional causality hypothesis with an AR(4) model on either co-variate.

```{r}

# Preparing the data
Y <- CPI.dat
X <- SENSEX.dat
# Computing Lags
Y.lag1 <- stats::lag(Y)
X.lag1 <- stats::lag(X)
Y.lag2 <- stats::lag(Y, k=2)
X.lag2 <- stats::lag(X, k=2)
Y.lag3 <- stats::lag(Y, k=3)
X.lag3 <- stats::lag(X, k=3)
Y.lag4 <- stats::lag(Y, k=4)
X.lag4 <- stats::lag(X, k=4)

# Fixing the lags
X <- X[1:(length(X)-5)]
X.lag1 <- X.lag1[2:(length(X.lag1)-4)]
X.lag2 <- X.lag2[3:(length(X.lag2)-3)]
X.lag3 <- X.lag3[4:(length(X.lag3)-2)]
X.lag4 <- X.lag4[5:(length(X.lag4)-1)]

Y <- Y[1:(length(Y)-5)]
Y.lag1 <- Y.lag1[2:(length(Y.lag1)-4)]
Y.lag2 <- Y.lag2[3:(length(Y.lag2)-3)]
Y.lag3 <- Y.lag3[4:(length(Y.lag3)-2)]
Y.lag4 <- Y.lag4[5:(length(Y.lag4)-1)]

## Direction 1 
m1 <- lm(Y~X+Y.lag1+Y.lag2+Y.lag3+Y.lag4)
summary(m1)

# Directrion 2
m2 <- lm(X~Y+X.lag1+X.lag2+X.lag3+X.lag4)
summary(m2)
```

Residuals plot for this:\

```{r}
# For Direction - 1
plot(m1$residuals)
adf.test(m1$residuals)

# For Direction - 2
plot(m2$residuals)
adf.test(m2$residuals)
```

```{r}

## Direction 1 
m1 <- lm(Y~X+X.lag1+Y.lag1+Y.lag2+Y.lag3+Y.lag4)
summary(m1)

## Directrion 2
m2 <- lm(X~Y+Y.lag1+X.lag1+X.lag2+X.lag3+X.lag4)
summary(m2)


```

```{r}

# For Direction - 1
plot(m1$residuals, ylab = "Residuals", main = "CPI on SENSEX")
adf.test(m1$residuals)

# For Direction - 2
plot(m2$residuals, ylab = "Residuals", main = "SENSEX on CPI")
adf.test(m2$residuals)
```

SWAP Model

```{r}
# Observing the data
plot(X,Y)
plot(X,Y.lag1)
plot(X,Y.lag2)
plot(X,Y.lag3)
plot(X,Y.lag4)
```

```{r}
# Observing the data
plot(Y,X)
plot(Y,X.lag1)
plot(Y,X.lag2)
plot(Y,X.lag3)
plot(Y,X.lag4)
```
